Bootstrap Confidence Intervals for Regression Discontinuity Designs

### 2016-02-14 main_paper.tex
Add argmin operator to improve formatting and spacing
This commit adds 'argmin' and 'argmax' operators. It also creates a new file, 'macros.tex',
for other operator definitions.
Section 2, not quite happy with the result because it is not accessible to someone not used to the literature. I would like to know your opinion in terms of: (i) Do you think it would be ok to have  just a general description of what they do without having to go through the equations I have included now? I tried to make it as free of their equations as possible, but it still feels very terse. (ii) Do I need to get into

\documentclass[12pt,fleqn]{article}
\input{VERSION}
\input{tex/setup}

\title{Bootstrap Confidence Intervals for Regression Discontinuity Designs}
\author{Ot\'avio Bartalotti \and Gray Calhoun \and Yang He\thanks{All authors: 
Department of Economics, Iowa State University. 260 Heady Hall, Ames, IA  50011.
Bartalotti: Email: bartalot@iastate.edu. Calhoun: gcalhoun@iastate.edu.
He: yanghe@iastate.edu.}}

@@

\section{Background}\label{background}
In the typical sharp RD setting, a researcher wishes to estimate the local 
causal effect of treatment at a given threshold. The running variable, $X_{i}$, 
determines treatment assignment.  Given a known threshold, $\bar{x}$, set to 
zero without loss of generality, a unit receives treatment if $X_{i} \geq 0$ or 
does not receive treatment if $X_{i} < 0$. Let $Y_{i}(1)$ and $Y_{i}(0)$ denote 
the potential outcomes for unit $i$ given it receives treatment and
 in the absence of treatment, respectively. Hence, the observed sample is 
 comprised of the running variable, $X_{i}$, and
 \begin{align}
  Y_{i}=Y_{i}(0) \mathbbm{1}\{X_{i}<0\}+Y_{i}(1) \mathbbm{1}\{X_{i} \geq 0\}
 \end{align}
where $\mathbbm{1}\{ \cdot\}$ denotes the indicator function. For convenience, 
define
 \begin{align}
  \mu(x)= \mathbbm{E}[Y_{i}|X_{i}=x]
 \end{align}
Also, let $\mu^{(\eta)}(x)=\frac{d^{\eta}\mu(x)}{dx^{\eta}}$ be the $\eta^{th}$ 
derivative of the unknown regression function and define 
$\mu^{(\eta)}_{+}=lim_{x \rightarrow 0^{+}}\mu^{(\eta)}(x)$ and 
$\mu^{(\eta)}_{-}=lim_{x \rightarrow 0^{-}}\mu^{(\eta)}(x)$.
In most cases the population parameter of interest is 
$\tau=\mathbbm{E}[Y(1)-Y(0)|X=\bar{x}]$ (i.e., the average treatment effect 
at the threshold). Under continuity and smoothness conditions on both the 
conditional distribution of $X_i$ and the first moments of $Y(0)$ and $Y(1)$ at
the cutoff, $\tau$ is nonparametrically identifiable (Hahn, Todd and Van der 
Klaauw, 2001) by:
\begin{align}
 \tau&= \mu_{+}- \mu_{-} \nonumber \\ 
&\text{ where } \mu_{+}=lim_{x \rightarrow 0^{+}}\mu(x),\text{ and } 
\mu_{-}=lim_{x \rightarrow 0^{-}}\mu(x)
\end{align}

Naturally, the estimation of $\tau$ in RD designs focuses on the 
problem of approximating $\mathbbm{E}[Y(1)|X=x]$ and $\mathbbm{E}[Y(0)|X=x]$ 
near the cutoff. Due to its desirable properties when estimating regression 
functions at boundary points, we consider the popular approach of fitting 
separate kernel-weighted local linear regressions in neighborhoods on both 
sides of the threshold.\footnote{See \cite{HTV2001}, \cite{Porter03} or
\cite{FanGijbels92} for discussions of the properties of local polynomial 
regressions for boundary problems.The results presented here are valid for
higher order polynomials and discontinuities at derivatives, like ``Kink RDD''
in Card, Lee and Pei (2009). We focus on the linear case for ease in exposition.}
For the local linear model, we use the following estimator as described in 
Calonico, Cattaneo and Titiunik (2014)\footnote{Throughout the paper we follow the
notation on Calonico, Cattaneo and Titiunik (2014) very closely.}
\begin{align*}
 \hat{\tau}(h_{n})&=\hat {\mu}_{+}(h_{n}) -\hat{\mu}_{-}(h_{n})\\
(\hat {\mu}_{+}(h_{n}),\hat {\mu}^{(1)}_{+}(h_{n}))'&= argmin_{b_{0},b_{1}} 
\sum_{i=1}^{N}\mathbbm{1}\{X_{i} \geq 0\}(Y_{i}-b_{0}-X_{i}b_{1})^{2} \cdot K_{h}(X_{i})\\
(\hat {\mu}_{-}(h_{n}),\hat {\mu}^{(1)}_{-}(h_{n}))' &= argmin_{b_{0},b_{1}}
\sum_{i=1}^{N}\mathbbm{1}\{X_{i} < 0\}(Y_{i}-b_{0}-X_{i}b_{1})^{2}\cdot K_{h}(X_{i})
\end{align*}
where $K_{h}(x_{i}) = K\left(\frac{x_{i}}{h}\right)\frac{1}{h}$.

>>>>

\documentclass[12pt,fleqn]{article}
\input{VERSION}
\input{tex/setup}
\input{tex/macros}

\author{Ot\'avio Bartalotti \and Gray Calhoun \and Yang He\thanks{All authors: 
Department of Economics, Iowa State University. 260 Heady Hall, Ames, IA  50011.
Bartalotti: Email: bartalot@iastate.edu. Calhoun: gcalhoun@iastate.edu.
He: yanghe@iastate.edu.}}

@@

\section{Background}\label{background}

In the typical sharp RD setting, a researcher wishes to estimate the local 
causal effect of treatment at a given threshold. The running variable, $X_{i}$, 
determines treatment assignment.  Given a known threshold, $\bar{x}$, set to 
zero without loss of generality, a unit receives treatment if $X_{i} \geq 0$ or 
does not receive treatment if $X_{i} < 0$. Let $Y_{i}(1)$ and $Y_{i}(0)$ denote 
the potential outcomes for unit $i$ given it receives treatment and
 in the absence of treatment, respectively. Hence, the observed sample is 
 comprised of the running variable, $X_{i}$, and
 \begin{align}
  Y_{i}=Y_{i}(0) \mathbbm{1}\{X_{i}<0\}+Y_{i}(1) \mathbbm{1}\{X_{i} \geq 0\}
 \end{align}
where $\mathbbm{1}\{ \cdot\}$ denotes the indicator function. For convenience, 
define
 \begin{align}
  \mu(x)= \mathbbm{E}[Y_{i}|X_{i}=x]
 \end{align}
Also, let $\mu^{(\eta)}(x)=\frac{d^{\eta}\mu(x)}{dx^{\eta}}$ be the $\eta^{th}$ 
derivative of the unknown regression function and define 
$\mu^{(\eta)}_{+}=lim_{x \rightarrow 0^{+}}\mu^{(\eta)}(x)$ and 
$\mu^{(\eta)}_{-}=lim_{x \rightarrow 0^{-}}\mu^{(\eta)}(x)$.
In most cases the population parameter of interest is 
$\tau=\mathbbm{E}[Y(1)-Y(0)|X=\bar{x}]$ (i.e., the average treatment effect 
at the threshold). Under continuity and smoothness conditions on both the 
conditional distribution of $X_i$ and the first moments of $Y(0)$ and $Y(1)$ at
the cutoff, $\tau$ is nonparametrically identifiable (Hahn, Todd and Van der 
Klaauw, 2001) by:
\begin{align}
 \tau&= \mu_{+}- \mu_{-} \nonumber \\ 
&\text{ where } \mu_{+}=lim_{x \rightarrow 0^{+}}\mu(x),\text{ and } 
\mu_{-}=lim_{x \rightarrow 0^{-}}\mu(x)
\end{align}

Naturally, the estimation of $\tau$ in RD designs focuses on the 
problem of approximating $\mathbbm{E}[Y(1)|X=x]$ and $\mathbbm{E}[Y(0)|X=x]$ 
near the cutoff. Due to its desirable properties when estimating regression 
functions at boundary points, we consider the popular approach of fitting 
separate kernel-weighted local linear regressions in neighborhoods on both 
sides of the threshold.\footnote{See \cite{HTV2001}, \cite{Porter03} or
\cite{FanGijbels92} for discussions of the properties of local polynomial 
regressions for boundary problems.The results presented here are valid for
higher order polynomials and discontinuities at derivatives, like ``Kink RDD''
in Card, Lee and Pei (2009). We focus on the linear case for ease in exposition.}
For the local linear model, we use the following estimator as described in 
\cite{calonico2014},\footnote{Throughout the paper we follow the
notation on \cite{calonico2014} very closely.}
\begin{align*}
 \hat{\tau}(h_{n})&=\hat {\mu}_{+}(h_{n}) -\hat{\mu}_{-}(h_{n})\\
(\hat {\mu}_{+}(h_{n}),\hat {\mu}^{(1)}_{+}(h_{n}))'&= argmin_{b_{0},b_{1}} 
\sum_{i=1}^{N}\mathbbm{1}\{X_{i} \geq 0\}(Y_{i}-b_{0}-X_{i}b_{1})^{2} \cdot K_{h}(X_{i})\\
(\hat {\mu}_{-}(h_{n}),\hat {\mu}^{(1)}_{-}(h_{n}))' &= argmin_{b_{0},b_{1}}
\sum_{i=1}^{N}\mathbbm{1}\{X_{i} < 0\}(Y_{i}-b_{0}-X_{i}b_{1})^{2}\cdot K_{h}(X_{i})
\end{align*}
where $K_{h}(x_{i}) = K\left(\frac{x_{i}}{h}\right)\frac{1}{h}$.

As shown by \cite{HTV2001}, \cite{Porter03} and expanded by \cite{calonico2014} 
identification and inference procedures can be developed based on the following 
assumptions\footnote{The assumptions below are the same as presented by \cite{calonico2014}.}
\begin{assumption}\label{A1}
 For some $\kappa_{0} > 0$, the following holds in the neighborhood 
 $(-\kappa_{0},\kappa_{0})$ around the cutoff $ \bar{x}= 0$:
 \begin{enumerate}
  \item $E[\left.Y_{i}^{4}\right|X_{i}=x]$ is bounded and the density of $X$, 
  $f(x)$, is continuous and bounded away from zero.
  \item $\mu_{+}(x)=\mathbbm{E}[\left.Y_{i}(1)\right|X_{i}=x]$ and $\mu_{-}(x)=
  \mathbbm{E}[\left.Y_{i}(0)\right|X_{i}=x]$ are $S$ times continuously 
  differentiable.
  \item $\sigma^{2}_{+}(x)=\mathbbm{V}[\left.Y_{i}(1)\right|X_{i}=x]$ and 
  $\sigma^{2}_{-}(x)=\mathbbm{V}[\left.Y_{i}(0)\right|X_{i}=x]$ are continuous 
  and bounded away from zero.
 \end{enumerate}
\end{assumption}

The conditions in Assumption \ref{A1} are the usual smoothness and existence 
conditions in which the RD literature relies. Its second part is important for 
the characterization of the leading bias term that will be the focus of the bias
correction procedures described in \cite{calonico2014} and in the novel
 bootstrap procedure proposed in Section \ref{boot}.

The second set of assumptions regards the possible kernel weights used in RD.

\begin{assumption}\label{A2}
 For some $\kappa>0$, the kernel function $k(\cdot):[0, \kappa] \rightarrow 
 \mathbbm{R}$ is bounded and nonnegative, zero outside its support, and positive
 and continuous on $(0, \kappa)$.
\end{assumption}

Assumption \ref{A2} supports the kernels most commonly used in applications, in 
particular the broadly used uniform kernel $k(u)=\mathbbm{1}[0\leq u\leq 1]$, 
which simplifies to the use of a local linear OLS on both sides of the treshold.
Since the objective of this paper is to present a simple bootstrap procedure to 
obtain robust confidence intervals in the context of RD designs that can be 
easily implemented by practitioners, during the remainder of the paper we focus 
on estimators that use the rectangular kernel. The use of different weighting 
structures implied by different kernels complicate the bootstrap algorithm in
meaningful ways and such analysis is left to future work.

\cite{calonico2014} point out that the conventional approaches to construct 
confidence intervals for $\tau$ using the local linear estimator rely on a 
large-sample approximation for the standardized $t$-statistic that is valid only
if the bandwidth shrink fast enough to eliminate the leading bias term 
contribution to the approximation. Hence, if $nh_{n}^{5}\rightarrow0$ and 
$nh_{n}\rightarrow \infty$, then
 \begin{align}
  T(h_{n})=\frac{\hat{\tau}(h_{n})-\tau}{\sqrt{V(h_{n})}}\rightarrow_{d}N(0,1),
  \text{        }V(h_{n})=\mathbbm{V}[\left.\tau(h_{n})\right|\chi_{n}],
  \text{        }\chi_{n}=[X_{1}, \dots X_{n}]^{\prime}
 \end{align}
 The conditional variance $V(h_{n})$ depends on the $\sigma^{2}_{+}=
 lim_{x \rightarrow 0^{+}}\sigma^{2}_{+}(x)$ and $\sigma^{2}_{-}=
 lim_{x \rightarrow 0^{-}}\sigma^{2}_{-}(x)$ as well as $h_{n}$ and known 
 quantities that depend on the kernel and order of polynomial used in the 
 estimation. Under these conditions we could use a conventional confidence 
 interval for $\tau$ given by
 \begin{align}
  I(h_{n})=\left[\hat{\tau}(h_{n})\pm \Phi^{-1}_{1-\frac{\alpha}{2}}\sqrt{V(h_{n})} \right]
 \end{align}
with $\Phi^{-1}_{1-\frac{\alpha}{2}}$ being the $\alpha$-quantile of the 
standard normal distribution. However, most approaches to select the bandwidth, 
$h_{n}$ in the literature, including the widely used optimal-AMSE bandwidth 
selector proposed by \cite{IK} lead to bandwidth choices that are ``too large'' 
since they do not satisfy the condition $nh_{n}^{5}\rightarrow0$, leading to a 
first-order bias in the distributional approximation used to construct the 
confidence intervals.

\cite{calonico2014} resolve the problem by deriving the analytical form of the 
first-order bias and correcting it directly. To obtain correct coverage the 
confidence interval needs not only to be recentered to correct the bias, but 
also rescaled to allow for the additional variability introduced by the bias 
correction. Their approach rely on obtaining a bias-corrected  estimator, 
$\hat{\tau}^{bc}$ and its adjusted variance. It is important first to note that 
the first-order bias term itself will depend on the bandwidth $h_{n}$. The 
approximate bias term can be described as:
\begin{align*}
     E[\hat{\tau}(h_{n})|\chi_{n}]-\tau=& h_{n}^{2}\mathsf{B}(h_{n})\{1+o_{p}(1)\}\\
     \mathsf{B}(h_{n})=&\frac{\mu_{+}^{(2)}}{2!}\mathfrak{B}_{+}(h_{n})-\frac{\mu_{-}^{(2)}}{2!}\mathfrak{B}_{-}(h_{n})
   \end{align*}
where $\mathfrak{B}_{+}(h_{n})$ and $\mathfrak{B}_{-}(h_{n})$ are observed 
quantities that depend on the data, kernel and $h_{n}$. The plug-in 
bias-corrected estimator then requires estimates for the second derivatives of 
the conditional mean from above and below the cutoff, $\mu_{+}^{(2)}$ and 
$\mu_{-}^{(2)}$. \cite{calonico2014} propose using a conventional local quadratic
estimator, i.e. one order higher than the polynomial used to obtain $\hat{\tau}$,
using a (potentially) different pilot bandwidth $b_{n}$.
\begin{align*}
     \hat{\tau}^{bc}(h_{n}, b_{n})=& \hat{\tau}-h_{n}^{2}\hat{\mathsf{B}}(h_{n},b_{n})\\
     \hat{\mathsf{B}}(h_{n})=&\frac{\hat{\mu}_{+}^{(2)}(b_{n})}{2!}
     \mathfrak{B}_{+}(h_{n})-\frac{\hat{\mu}_{-}^{(2)}(b_{n})}{2!}\mathfrak{B}_{-}(h_{n})
   \end{align*}
By incorporating the contribution of both $\hat{\tau}(h_{n})$ and 
$\hat{\mathsf{B}}(h_{n},b_{n})$ to the asymptotic distribution of the estimator,
\cite{calonico2014} obtain a robust confidence interval with a different 
asymptotic variance in general. So under Assumptions\ref{A1} and \ref{A2}, if 
$n\min\{h_{n}^{5}, b_{n}^{5}\}max\{h_{n}^{2}, b_{n}^{2}\}\rightarrow 0$ and 
$n\min\{h_{n}, b_{n}\}\rightarrow \infty$ and $\kappa \max\{h_{n},b_{n}\}< 
\kappa_{0}$, they show that
\begin{align}
  T^{rbc}(h_{n}, b_{n})=\frac{\hat{\tau}^{bc}(h_{n}, b_{n})-\tau}{\sqrt{V^{bc}(h_{n}, b_{n})}}\rightarrow_{d}N(0,1),
  \text{        }V^{bc}(h_{n}, b_{n})=V(h_{n})+C^{bc}(h_{n}, b_{n})\\
 \end{align}
where the additional term $C^{bc}(h_{n}, b_{n})$ depends on the (asymptotic) 
variability of the bias estimate used for correction as well as its correlation 
with the original RD estimator $\hat{\tau}(h_{n})$.\footnote{The details of the 
formulas are currently ommitted from this paper since their form is not relevant
to the bootstrap procedure innovation proposed. The details can be found at 
\cite{calonico2014} Appendix}  Under these conditions we can construct a valid 
confidence interval for $\tau$ given by
 \begin{align}
  I^{rbc}(h_{n}, b_{n})=\left[\left(\hat{\tau}(h_{n})-h_{n}^{2}\hat{\mathsf{B}}
  (h_{n},b_{n})\right)\pm \Phi^{-1}_{1-\frac{\alpha}{2}}\sqrt{V(h_{n})+C^{bc}
  (h_{n}, b_{n})} \right]
 \end{align}

This approach is shown by the authors to significantly coverage of the CIs 
constructed and provides practitioners with a new toolset to perform inference 
that is more robust to the choice of bandwidth. We build upon the insight 
provided by \cite{calonico2014} bias-corrected estimator and propose a simple 
bootstrap procedure that can directly construct the robust CIs without requiring
the derivation of analytical formulas and direct estimators for the bias, 
variance and covariance terms, while relying on the same first-order bias 
correction approximation proposed by that paper.
